### 2024-03-02 script/CoverLetter.Rmd
c

---
output: pdf_document
---

\begin{flushleft}
Dr. Aaron Ellison \newline
Executive Editor \newline
Methods in Ecology and Evolution \newline
\end{flushleft}

March 17^th^, 2024

Attached is our standard article submission "*Protocol to increase accuracy and fidelity of pollen meta-genomic barcoding using Angiosperms353: a case study using pollen loads from wild bumble bees*".

In this paper, we combine Hyb-Seq to generate *de novo* sequence databases, informed by museum records, and species distribution modelling. 
Further we use the results of phenological modelling to interpret our sequence alignment results to increase the likelihood of accurate results.

Given the rising popularity of metabarcoding and metagenomics, especially for projects over large spatial domains and with multiple seasonal time points, we believe that our method provides a framework for meaningfully assessing and contextualizing results. 
We have observed that many studies utilizing metabacoding generate spurious results, which are precluded due to spatial or temporal incongruities with the focal animal species and plants. 
However, a simple and intuitive framework for programmatically assessing and rectifying these issues are lacking.

We believe that the use of the Angiosperms353 probes was novel at the inception of the study, and that several manuscripts utilizing them will be forthcoming to a variety of journals.

Our submission contains X supplementary materials, X figures and X tables. It unfortunately is currently at roughly 9,300 words. 
However *circa* 900 words are dedicated to a plain text overview of the methods, a section which we believe elucidates the integration of the various approaches, and clarifies adoption by subsequent users. 
Considerable length is dedicated to the methods, which contain five years of field work, and similarly thorough microscopy, and molecular work; but which we have cut by nearly 500 words.

We have no conflicts of interest to disclose. Additionally, this work - nor any part of it - has been submitted to any other journal.

Please address all correspondence regarding this manuscript to me at rbenkendorf@chicagobotanic.org

Thank you for considering this work.

\begin{flushleft}
Reed Clark Benkendorf \newline
Senior (Field) Botanist \& Senior Spatial Data Specialist  \newline
The Chicago Botanic Garden  \newline
rbenkendorf@chicagobotanic.org
\end{flushleft}

\begin{flushleft}
Jeremie B. Fant, Ph.D \newline
Medard and Elizabeth Welch Director of Conservation \newline
The Chicago Botanic Garden  \newline
jfant@chicagobotanic.org
\end{flushleft}