### 2024-03-29 script/StatementOnInclusion.Rmd

---
title: "Statement On Inclusion"
output:
  word_document: default
  pdf_document: default
---

Our studies focal area was within the Southern Rockies Mountains of Western North America and drew on natural history data collected throughout the region. 
Two of our authors are *staff?affiliates?* at the research station which novel data was collected at. 
Prior to matriculation our corresponding author worked at the largest institute for applied ecology in the Rocky Mountains, and has since worked as a contractor for the largest land management agency in the United States and is both based in and works in this geographic area.