### 2024-03-30 script/SpecimensSequencedTable.Rmd
ideally simplyfing repo for code submission

---
output: pdf_document
classoption: landscape
geometry: "left=1cm,right=1cm,top=1cm,bottom=1cm"
header-includes:
  - \usepackage{titlesec}
  - \usepackage{caption}
  - \pagenumbering{gobble}
  - \titlespacing{\title}{0pt}{\parskip}{-\parskip}
---

\captionsetup[table]{labelformat=empty}
\vspace{-25truemm}
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, warning = F, message = F)
```

```{r Required Libraries, message=F, warning=F, results='hide'}
library(knitr)
library(kableExtra)
library(tidyverse)
```

```{r}
dat <- read.csv('../data/Reference_library_for_appendix.csv') %>% 
  arrange(Family, Taxon) %>% 
  unite('Locality', Country:County, sep = ', ') %>% 
  mutate(Locality = str_replace(Locality, 'U[.]S[.]A[.]', 'USA')) %>% 
  select(-Coordinates) %>% 
  rename(Herbarium = 'Herb.', Preservation = 'Pres.')

kable(
  dat, 
  caption = 'Samples Used to Supplement the Sequence Reference Library',
  scaledown = T, 
  booktabs = TRUE, 
  format  = "latex") %>% 
  kableExtra::kable_styling(latex_options="scale_down")

```